---
title: "Exercises 1.4-1.10 <br> solution"
author: "Niels Richard Hansen"
date: "September 4, 2019"
output:
  xaringan::moon_reader:
    css: ["default", "science.css"]
    nature:
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:10'
      navigation:
        scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(purrr)
library(microbenchmark)
library(Rcpp)
library(ggplot2)
library("profvis")
library(MASS)
library(parallel)
library(numDeriv)
setwd("~/Documents/Statistics - KU/Computational Statistics/Assignments")
set.seed(2)
```

Sample from the probability distribution on $[0;\infty[$ with a density
$$f(y) \propto \prod_{i=1}^{100}\exp(yz_ix_i-e^{yx_i}), \text{ } y\geq 0$$
Find a Gaussian envelope of $f$ and implement rejection sampling from the distribution with density $f$ using this envelope.
---
#### How to evaluate the target density
```{r}
pois <- read.table("Poisson.csv", sep = ",", header = TRUE)
zx <- pois$z*pois$x
x <- pois$x
```

Let's look at a couple of ways to compute this product.

```{r}
# Unnormalized

target_sapply <- function(y){
  sapply(y, function(yy) prod(exp(yy*zx - exp(x*yy))))
}

target_sum <- function(y){
  sapply(y, function(yy) exp( sum(yy*zx - exp(x*yy)) ))
}

target_outer <- function(y){
  exp( rowSums(outer(y, zx) - exp(outer(y, x))))
}
```
---
We see that these densities need to be normalized.

```{r}
value_sapply <- integrate(target_sapply, 0, Inf)$value
value_sum <- integrate(target_sum, 0, Inf)$value
value_outer <- integrate(target_outer, 0, Inf)$value
```
And we can compute:

```{r}
target_sapply <- function(y){
  sapply(y, function(yy) prod(exp(yy*zx - exp(x*yy)))) / value_sapply
}
target_sum <- function(y){
  sapply(y, function(yy) exp( sum(yy*zx - exp(x*yy)) )) / value_sum
}
target_outer <- function(y){
  exp( rowSums(outer(y, zx) - exp(outer(y, x))))/ value_outer
}
```
---
```{r}
grid <-seq(0, 0.5, 0.0001)
```
---
#### See which method is the best.

```{r, message = FALSE}
p <- microbenchmark(times = 200, target_sum(grid), target_sapply(grid), target_outer(grid))
autoplot(p) + geom_jitter(aes(color = expr), alpha = 0.4) + aes(fill = I("gray"))+
  scale_color_manual(values = c("#354e75", "#884091", "#c91246", "black")) + theme_bw() +
  theme(legend.position = "none")
```
---
Since we need to find a Gaussian envelope

```{r}
mean <- integrate(function(v) target_sapply(v)*v, 0, Inf)$value
sd <- sqrt(integrate(function(v) target_sapply(v)*v^2, 0, Inf)$value - mean^2)
```
---
Let see what the target density
```{r}
plot(grid,target_outer(grid), type = "l", lwd = 4, col = "#325b96", ylab = "Density", xlab = "")
```
---
#### what we need to minimize is

```{r}
plot(grid, dnorm(grid, mean, sd)/target_outer(grid), lwd =1, col = "#325b96" )
```
---
We get
```{r}
a <- optimize(function(z) dnorm(z, mean, sd)/target_sapply(z), c(0, 0.1))
a <- a$objective
```

So we get that our alpha is `r round(a,4)`. With the chosen alpha, we can replot the envelope and see what we get

```{r}

plot(grid, a*target_sapply(grid), type = "l", lwd = 3,  col = "#325b96",ylim  = c(0,7))
lines(grid, dnorm(grid, mean, sd),  col = "#ee961d", lwd =3)
```

---
We see that this is not good enough. with trial and error we found out that by increasing the standard deviation by 9% we get a much better estimate:
```{r}
a <- optimize(function(z) dnorm(z, mean, sd*1.09)/target_sapply(z), c(0, 0.1))
a <- a$objective
```

So we get that our alpha is `r round(a,4)`. With the chosen alpha, we can replot the envelope and see what we get

```{r}
plot(grid, a*target_sapply(grid), type = "l", lwd = 3,  col = "#325b96")
lines(grid, dnorm(grid, mean, sd),  col = "#ee961d", lwd =3)
```

We also tried to use the t-distribution to get an envelope:
---
```{r}
at <- optimize(function(z) dt((z - 0.24)/(0.0554*1.055), 61.83)/(0.0554*1.055) / target_sapply(z), c(0, 0.4))$objective
#t distribution
plot(grid, at*target_sapply(grid), type = "l", lwd = 3,  col = "#325b96")
lines(grid, dt((grid - 0.24)/(0.0554*1.055), 61.83)/(0.0554*1.055),  col = "#ee961d", lwd =3)
```

We now look at two different implementations
```{r}
proposal_factory <- function(target_dens, proposal_dens, proposal_sim, a)
{
  list( tdens = target_dens, pdens = proposal_dens, sim = proposal_sim, alpha = a)
}
propn <- proposal_factory(target_outer, function(x) dnorm(x, 0.2388885, 0.061235426), function(x) rnorm(x, 0.2388885, 0.06123542), 0.9163125)
propt <- proposal_factory(  
  target_outer,
  function(x) dt((x - 0.24)/(0.0554*1.055), 61.83)/(0.0554*1.055),
  function(x) rt(x, 61.83)*0.0554*1.055 + 0.24,
  0.95
)
```

# Setja inn kóða með fyrri implementation.

